---
title: "Visualizing Linkage Disequlibrium"
author: "Bruce Cochrane"
date: "October 6, 2016"
output: html_document
---

### Load the data

Processed data have been stored as a list of two (CEU and YRI).  We need to load them and go from there
```{r}
load("./ProcessedData/lct.rda")
```

And load functions (need to automate this)

```{r}
source("./R/VcfFunctions.R")
```

Note that with names added, we can work directly from the list:

```{r}
names(lct)
```

### Make a distance object

```{r}
yri.r <-LDscan(lct$yri.vcf)
ceu.r <-LDscan(lct$ceu.vcf)
```

Note that the above could be done with sapply

```{r}
dat.r <-sapply(lct,LDscan)
```

### Plotting with pegas

```{r}
dat.r.str <-sapply(dat.r,lblstrip)
```


```{r}
par(mfrow=c(2,1))
sapply(dat.r.str,LDmap,scale.legend=3.5,cex=.4)
par(mfrow=c(1,1))
```

### plotting with LDheatmap

```{r}
library(LDheatmap)
```

Convert distance objects to matrices

```{r}
dat.r.mat <-sapply(dat.r,as.matrix)
```

Now take a shot

```{r}
sapply(dat.r.mat,LDheatmap,color=heat.colors(20))
```

Overall, this seems like the better solution - with work, will have better display (i. e. map by position.)

Thinking about tag snps

```{r}
range(dat.r.mat)
```

